---
title: "HW 4"
author: "Anish Palley, ap65675, https://github.com/anishpalley/SDS-315-HW4"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: '3'
    df_print: paged
  pdf_document: default
---

```{r global options, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, # show the code
                      eval = TRUE, # run the code
                      warning = FALSE, # show warning messages
                      message = FALSE, # show basic messages
                      fig.align = "center"
                      )
#Uploading necessary packages
library(tidyverse)
library(ggplot2)
library(rvest)
library(dplyr)
library(plotly)
library(mosaic)
```

### Problem 1

Null Hypothesis: The null hypothesis is that over the long run, securities trades from the Iron Bank are flagged at the same 2.4% baseline rate as that of other traders.

Test statistic: The test statistic is the number of trades flagged by the SEC's detection algorithm.

Probability distribution: Shown below, the probability distribution uses 100,000 samples generated using Monte Carlo simulation assuming that the Iron Bank trades are flagged at the 2.4% baseline rate.

```{r}
sim_flags = do(100000)*nflip(n=2021, prob=.024)

ggplot(sim_flags) + geom_histogram(aes(x=nflip), binwidth = 1, fill = 'skyblue', color = 'black')+
  labs(
    title = "Probability distribution of the number of trades flagged assuming the 2.4% baseline rate",
    x = "Number of trades flagged",
    y = "Count"
  )

flags_p_value <- sum(sim_flags >= 70)/100000
flags_p_value
```
P-value: The p-value shown above varies every run due to Monte Carlo variance, but it is around 0.002 and always below 0.05 which suggests that the observed number of trade flags, 70, is unlikely to have occurred by random chance. 

Conclusion: Since the p-value is less than 0.05, we reject the null hypothesis and have sufficient evidence to indicate that over the long run, securities trades from the Iron Bank are not flagged at the same 2.4% baseline rate as that of other traders, and that Iron Bank may be trading in violation of "insider trading" laws and should be further investigated.

### Problem 2

Null Hypothesis: The null hypothesis is that Gourmet Bites is consistent with the other restaurants in the city which are cited for health code violations at the same 3% baseline.

Test statistic: The test statistic is the number of health code violations reported

Probability distribution:Shown below, the probability distribution uses 100,000 samples generated using Monte Carlo simulation assuming that the Gourmet Bites is consistent with out restaurants in the city with health code violations at the 3% baseline.

```{r}
sim_violations = do(100000)*nflip(n=50, prob=.03)
ggplot(sim_violations) + geom_histogram(aes(x=nflip), binwidth = 1, fill = 'skyblue', color = 'black')+
  labs(
    title = "Probability distribution of the number of health code violations assuming the 3% baseline rate",
    x = "Number of health code violations",
    y = "Count"
  )
violations_p_value <- sum(sim_violations>=8)/100000
violations_p_value
```
P-value: The p-value shown above varies every run due to Monte Carlo variance, but it is around 0.00011 and always below 0.05 which suggests that the observed number of health code violations, 8, is unlikely to have occured by random chance. 

Conclusion: Since the p-value is less than 0.05, we reject the null hypothesis and have sufficient evidence to indicate that Gourmet Bites' rate of health code violations is higher than the citywide average of 3%, and that Gourmet Bites' needs to be more closely monitored for violations.

### Problem 3

Null Hypothesis, Ho: The null hypothesis, Ho, is that the distribution of jurors empaneled by this judge is not significantly different from the county's population proportions.

Test statistic, T: The test statistic, T, is the normalized square of the differences between the expected and observed number of jurors per group also known as chi-squared.

Probability distribution, P(T|Ho): The probability distribution of the normalized square of the differences between the expected and observed number of jurors per group, P(T|Ho), assuming that the distribution of jurors empaneled by this judge is not significantly different from the county's population proportions as shown below.

```{r}
expected_distribution = c(Group_1 = 0.30, Group_2 = 0.25, Group_3 = 0.20, Group_4 = 0.15, Group_5 = 0.10)
observed_counts =  c(Group_1 = 85, Group_2 = 56, Group_3 = 59, Group_4 = 27, Group_5 = 13)
simulated_counts = rmultinom(1, 240, expected_distribution)

chi_squared_statistic = function(observed, expected) {
  sum((observed - expected)^2 / expected)
}

expected_counts <- 240*expected_distribution

num_simulations = 100000
chi2_sim = do(num_simulations)*{
  simulated_counts = rmultinom(1, 240, expected_distribution)
  this_chi2 = chi_squared_statistic(simulated_counts, expected_counts)
  c(chi2 = this_chi2) # return a vector with names and values
}

hist(chi2_sim$chi2, main = "Chi-squared Distribution of Jurors", xlab = "Chi-squared Value", col = "lightblue", border= 'black')



jurors_p_value <- sum(chi2_sim>=chi_squared_statistic(observed_counts, expected_counts))/100000
jurors_p_value
```
P-value: The p-value shown above varies every run due to Monte Carlo variance, but it is always below 0.05 which suggests that the observed difference in jurors in each group compared to the country's population proportions is unlikely to have occurred by random chance.

Conclusion: Since the p-value is less than 0.05, we reject the null hypothesis and have sufficient evidence to indicate that the distribution of jurors empaneled by this judge is significantly different from the county’s population proportions. This indicates that the judge should warrant further investigation for systematic bias and improper peremptory challenges. Other explanations could be that the sources used to create the jury pool may have been biased or systemic under representation and over representation of certain groups of people.

### Problem 4

#### Part A:
```{r}
brown_sentences <- readLines("brown_sentences.txt")

letter_frequency <- read_csv("letter_frequencies.csv")
calculate_chi_squared = function(sentence, freq_table) {
  
  # Ensure letter frequencies are normalized and sum to 1
  freq_table$Probability = freq_table$Probability / sum(freq_table$Probability)
  
  # Remove non-letters and convert to uppercase
  clean_sentence = gsub("[^A-Za-z]", "", sentence)
  clean_sentence = toupper(clean_sentence)
  
  # Count the occurrences of each letter in the sentence
  observed_counts = table(factor(strsplit(clean_sentence, "")[[1]], levels = freq_table$Letter))
  
  # Calculate expected counts
  total_letters = sum(observed_counts)
  expected_counts = total_letters * freq_table$Probability
  
  # Chi-squared statistic
  chi_squared_stat = sum((observed_counts - expected_counts)^2 / expected_counts)
  
  return(chi_squared_stat)
}

chi_squares <- c()

for (line in brown_sentences) {
  chi_squared_val <- calculate_chi_squared(line,letter_frequency)
  chi_squares <- append(chi_squares, chi_squared_val)
}

hist(chi_squares, main = "Chi-squared Distribution of letter frequencies", xlab = "Chi-squared Value", col = "lightblue", border= 'black')

```

#### Part B:
```{r}

sentences <- c(
  "She opened the book and started to read the first chapter, eagerly anticipating what might come next.",
  "Despite the heavy rain, they decided to go for a long walk in the park, crossing the main avenue by the fountain in the center.",
  "The museum’s new exhibit features ancient artifacts from various civilizations around the world.",
  "He carefully examined the document, looking for any clues that might help solve the mystery.",
  "The students gathered in the auditorium to listen to the guest speaker’s inspiring lecture.",
  "Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.",
  "The chef demonstrated how to prepare a delicious meal using only locally sourced ingredients, focusing mainly on some excellent dinner recipes from Spain.",
  "They watched the sunset from the hilltop, marveling at the beautiful array of colors in the sky.",
  "The committee reviewed the proposal and provided many points of useful feedback to improve the project’s effectiveness.",
  "Despite the challenges faced during the project, the team worked tirelessly to ensure its successful completion, resulting in a product that exceeded everyone’s expectations."
)
p_vals <- c()

for (words in sentences){
  chi_sq_val <- calculate_chi_squared(words,letter_frequency)
  p_val <- sum(chi_squares >= chi_sq_val)/length(chi_squares)
  p_vals <- append(p_vals,round(p_val,3))
}
print(data.frame(p_values = p_vals))


```
The table above shows that the 6th sentence is the most likely to have been produced by an LLM and watermarked to subtly adjust its frequency distribution over letters. The 6th sentence has the smallest p-value of around 0.009, although it varies every run due to Monte Carlo variance. The p-value is the probability of observing a chi-squared statistic as extreme or more extreme than the one calculated assuming all the sentences were human generated. The p-value being below 0.05 suggests that there is significant evidence against all the sentences being made by humans, favoring that one sentence (sentence 6) was generated with a different letter frequency distribution, which could be an LLM. However, this analysis uses Brown Corpus as the human writing baseline, so differences from the baseline could be due to writing style or time period even if the sentence was written by a human.


